---
title: "Conditional infit versus outfit under family-wise error control"
subtitle: "Dichotomous and polytomous items, B = 1000, Westfall-Young"
author: "Magnus Johansson"
date: today
---

## Scope

Every study in this paper decides on conditional **infit**. The parametric
bootstrap in `RMitemInfitCutoff()` simulates outfit at the same time and stores
it, and `RMitemInfit()` gained a `statistic` argument
(easyRasch2 > 1.1.0) that reports and tests it through the same
Westfall-Young step-down. So the comparison costs one extra call per bootstrap
and no extra simulation.

The question is narrow. Holding the decision rule fixed at
`correction = "fwer"`, `alpha = .05` and `B = 1000`, does outfit control the
family-wise error rate as infit does, and does it detect anything infit misses?

Outfit is the unweighted mean of squared standardised residuals, so it is
dominated by responses far from the item location. If it earns a place beside
infit anywhere, it is on an **off-target** item. That contrast is the spine of
the design: within each arm one misfitting item sits on the person mean and one
sits well above it, run as separate conditions rather than nested, so detection
is attributable to targeting rather than to the number of misfitting items.

### Both statistics come from the same bootstrap

Within a replication the data, the item parameters, the bootstrap draws and the
step-down are shared. Only the column read from `$results` differs
(`InfitMSQ` or `OutfitMSQ`). Every infit-versus-outfit comparison below is
therefore paired at the replication level, which is what makes the discordance
counts in the last section interpretable.

### Misfit calibrated on the infit scale, not on rho

Following the Method section, `rho` is not portable across item formats, so the
two arms are anchored to a comparable **infit** level rather than to a common
`rho`. The target is the range of the items that the real phq9 responses flag,
roughly 1.23 to 1.32.

A misfitting item far from the person mean injects less infit than an on-target
one at the same `rho`, so the anchoring has to hold at both targeting levels.
`calibration_rho_targeting.R` scans it, 80 replications at *n* = 600:

| Arm | rho | infit on-target | infit off-target | outfit on-target | outfit off-target |
|---|---|---|---|---|---|
| `dich20` | 0.50 | 1.376 | 1.303 | 1.743 | 2.028 |
| `dich20` | **0.60** | **1.304** | **1.243** | 1.604 | 1.735 |
| `dich20` | 0.65 | 1.265 | 1.216 | 1.499 | 1.615 |
| `dich20` | 0.75 | 1.202 | 1.158 | 1.362 | 1.446 |
| `phq9` | **0.85** | **1.271** | **1.264** | 1.314 | 1.388 |

`rho` = 0.60 and 0.85 are the selected values. They put mean injected infit at
1.274 and 1.268, matched across formats and inside the real flagged range.

Two things in that table are worth carrying into the results. Injected infit
falls as the misfitting item moves off target while injected **outfit rises**,
which is the asymmetry the study exists to price. And the same infit injection
produces a far larger outfit excess in the dichotomous arm (1.60 to 1.74)
than in phq9 (1.31 to 1.39). Neither is a prediction about detection, because
outfit's null variance is larger too, by roughly a factor of three in the
dichotomous arm. Whether the larger signal survives the larger noise is the
question.

The study's own first results section reports the achieved values, so the
anchoring is verifiable from the run rather than only from the scan. If either
arm drifts outside the 1.2 to 1.35 band, add a `rho` level and re-render:
`rho` is a per-arm constant here, so widening it means adding it to the grid
keys first.

### Two arms

| | `dich20` | `phq9` |
|---|---|---|
| Items | 20 dichotomous | 9 polytomous, CML from real responses |
| Locations | `runif(20, -2, 2)`, seed 20250727 (as in the dichotomous studies) | real, range 2.46 logits |
| Latent SD | 1.5 | 1.39 (estimated) |
| On-target misfit | V8, location 0.04 | q5, location -0.15 |
| Off-target misfit | V9, location 1.82 | q8, location 1.40 |
| Generator | `eRm::sim.xdim()` | `sim_partial_score()` |

Sample sizes are 250 and 600 for both arms. 600 is also the size of the real
phq9 sample the polytomous arm is derived from.

### Cost

Three conditions per arm (null, on-target, off-target), two sample sizes, two
arms, 250 replications. 3000 cells, each one bootstrap of 1000 iterations plus
three `RMitemInfit()` calls. Roughly five hours on ten cores. Cells are ordered
so the null conditions and the cheaper sample size finish first, since the null
carries the error-rate claim.

```{r setup}
suppressMessages({
  source("../R/helpers.R")   # package internals and shared helper functions
  library(easyRasch2)
  library(eRm)
  library(MASS)
  library(parallel)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
})

NS      <- c(250, 600)
B       <- 1000          # the decision rule under study
REPS    <- 250
ALPHA   <- 0.05
NCORES  <- 10
CHUNK   <- 100
RES     <- "infit_outfit.rds"

# This study needs the `statistic` argument, added to RMitemInfit() after
# easyRasch2 1.1.0. Checked here rather than five hours into the run.
if (!"statistic" %in% names(formals(easyRasch2::RMitemInfit))) {
  stop("Installed easyRasch2 (", packageVersion("easyRasch2"), ") has no ",
       "`statistic` argument in RMitemInfit(). Install the development ",
       "version before running this study.", call. = FALSE)
}
```

```{r runner}
# run_resumable() is defined in ../R/helpers.R
```

### Arm definitions

```{r arms}
# ---- dich20: the item locations used by the dichotomous studies -------------
set.seed(20250727)
D_LOCS <- runif(20, -2, 2)
D_ON   <- which.min(abs(D_LOCS))          # V8,  0.04 logits
D_OFF  <- which.max(D_LOCS)               # V9,  1.82 logits
D_SD   <- 1.5

# ---- phq9: CML estimates from the packaged real data ------------------------
ph <- load_phq9_arm()

ARMS <- list(
  dich20 = list(locs = D_LOCS, sd = D_SD, nm = paste0("V", 1:20),
                on = D_ON, off = D_OFF, rho = 0.60, thr = NULL),
  phq9   = list(locs = ph$locs, sd = ph$sd, nm = paste0("q", 1:9),
                on = 5L, off = 8L, rho = 0.85, thr = ph$thr)
)

bind_rows(lapply(names(ARMS), function(a) {
  x <- ARMS[[a]]
  data.frame(arm = a, items = length(x$locs), rho = x$rho,
             latent_sd = round(x$sd, 2),
             on_target = paste0(x$nm[x$on], " (", round(x$locs[x$on], 2), ")"),
             off_target = paste0(x$nm[x$off], " (", round(x$locs[x$off], 2), ")"))
}))
```

```{r simfun}
# One dataset. `target` is "none" (single dimension), "on" or "off": the named
# item is generated from a second dimension correlating `rho` with the first.
sim_arm <- function(arm, n, target, seed) {
  a <- ARMS[[arm]]
  ni <- length(a$locs)
  mis <- switch(target, none = integer(0), on = a$on, off = a$off)
  set.seed(seed)
  S <- a$sd^2 * matrix(c(1, a$rho, a$rho, 1), 2, 2)
  th <- MASS::mvrnorm(n, mu = c(0, 0), Sigma = S)

  if (is.null(a$thr)) {
    # dichotomous, via the same generator as the dichotomous studies
    W <- cbind(rep(1, ni), rep(0, ni))
    if (length(mis)) W[mis, ] <- c(0, 1)
    d <- as.data.frame(eRm::sim.xdim(persons = th, items = a$locs,
                                     Sigma = S, weightmat = W))
  } else {
    d <- sim_partial_score(a$thr, th[, 1])
    if (length(mis)) d[, mis] <- sim_partial_score(a$thr[mis], th[, 2])[, 1]
    d <- as.data.frame(d)
  }
  names(d) <- a$nm
  list(data = d, mis = if (length(mis)) mis else 0L)
}
```

### Run

One bootstrap per replication, read twice. `RMitemInfit()` is called once
without a cutoff for the observed statistics and once per statistic with
`p_value = TRUE`, all against the same `RMitemInfitCutoff()` object.

```{r run}
one_rep <- function(g) {
  out <- try({
    arm <- as.character(g$arm); target <- as.character(g$target)
    n <- g$n; rep <- g$rep
    ai <- match(arm, names(ARMS)); ti <- match(target, c("none", "on", "off"))
    seed_data <- 4e8 + 1e7 * ai + 1e6 * ti + 1e3 * n + rep

    sd_out <- sim_arm(arm, n, target, seed_data)
    d <- sd_out$data
    items <- names(d)

    co <- RMitemInfitCutoff(d, iterations = B, parallel = FALSE,
                            seed = seed_data + 5e8, verbose = FALSE)

    fi <- suppressWarnings(
      RMitemInfit(d, cutoff = co, p_value = TRUE, correction = "fwer",
                  alpha = ALPHA, statistic = "infit", output = "dataframe"))
    fo <- suppressWarnings(
      RMitemInfit(d, cutoff = co, p_value = TRUE, correction = "fwer",
                  alpha = ALPHA, statistic = "outfit", output = "dataframe"))

    i1 <- match(items, fi$Item); i2 <- match(items, fo$Item)
    data.frame(
      item = seq_along(items),
      mis = sd_out$mis,
      infit_msq  = fi$Infit_MSQ[i1],
      outfit_msq = fo$Outfit_MSQ[i2],
      p_infit    = fi$p_infit[i1],
      padj_infit = fi$padj_infit[i1],
      p_outfit   = fo$p_outfit[i2],
      padj_outfit = fo$padj_outfit[i2],
      iters = co$actual_iterations
    )
  }, silent = TRUE)
  if (inherits(out, "try-error")) NULL else out
}

# Null conditions first (they carry the error-rate claim), then on-target and
# off-target, cheaper n first within each block.
grid <- do.call(rbind, lapply(c("none", "on", "off"), function(tg) {
  expand.grid(rep = 1:REPS, n = NS, target = tg, arm = names(ARMS),
              stringsAsFactors = FALSE)[, c("arm", "target", "n", "rep")]
}))
grid <- grid[order(match(grid$target, c("none", "on", "off")), grid$n), ]
rownames(grid) <- NULL
cat("cells to run:", nrow(grid), "\n")

s <- run_resumable(RES, grid, c("arm", "target", "n", "rep"), one_rep,
                   extra = list(arms = lapply(ARMS, function(a)
                     a[c("locs", "sd", "nm", "on", "off", "rho")]),
                     B = B, alpha = ALPHA))

R <- do.call(rbind, lapply(seq_len(nrow(s$grid)), function(i) {
  r <- s$res[[i]]
  if (is.null(r)) return(NULL)
  g <- s$grid[i, c("arm", "target", "n", "rep"), drop = FALSE]
  rownames(g) <- NULL   # otherwise cbind() warns once per cell
  cbind(g, r)
})) |>
  mutate(is_misfit = mis > 0 & item == mis,
         loc = mapply(function(a, i) round(s$arms[[a]]$locs[i], 2), arm, item),
         f_in  = as.integer(padj_infit  < ALPHA),
         f_out = as.integer(padj_outfit < ALPHA),
         f_any = as.integer(f_in == 1 | f_out == 1))

cat(nrow(s$grid), "cells complete |", round(s$elapsed_min, 1), "min\n")
cat("replications with a failed cell:", sum(vapply(s$res, is.null, logical(1))), "\n")
```

## Achieved misfit, on both scales

Read this first. The two arms are meant to inject comparable misfit on the
infit scale, at roughly the level the real phq9 items reach. The outfit column
shows what the same injection does to the unweighted statistic, which is the
mechanism the rest of the study prices.

```{r calib}
R |>
  filter(is_misfit) |>
  group_by(arm, target, loc, n) |>
  summarise(infit = round(mean(infit_msq), 3),
            outfit = round(mean(outfit_msq), 3),
            sd_infit = round(sd(infit_msq), 3),
            sd_outfit = round(sd(outfit_msq), 3), .groups = "drop") |>
  arrange(arm, target, n)
```

Fitting items under the null, for reference:

```{r calib-null}
R |>
  filter(target == "none") |>
  group_by(arm, n) |>
  summarise(infit = round(mean(infit_msq), 3),
            outfit = round(mean(outfit_msq), 3),
            sd_infit = round(sd(infit_msq), 3),
            sd_outfit = round(sd(outfit_msq), 3), .groups = "drop")
```

## Family-wise error under a true model

The primary calibration result. A replication contributes one indicator per
rule: at least one item flagged when no item misfits. `union` is the rule an
analyst applies by reading both columns of a single table and flagging either,
which tests 2*k* hypotheses while each correction covers *k*.

```{r fwe}
R |>
  filter(target == "none") |>
  group_by(arm, n, rep) |>
  summarise(infit = as.integer(sum(f_in) > 0),
            outfit = as.integer(sum(f_out) > 0),
            union = as.integer(sum(f_any) > 0), .groups = "drop") |>
  group_by(arm, n) |>
  summarise(across(c(infit, outfit, union), \(x) round(100 * mean(x), 1)),
            reps = n(), .groups = "drop") |>
  mutate(mcse = round(100 * sqrt(0.05 * 0.95 / reps), 2))
```

Marginal calibration, before the step-down. This separates a miscalibrated
statistic from a miscalibrated correction: if the marginal rate is nominal and
the family-wise rate is not, the studentised-max step-down is what fails for
outfit, not outfit itself.

```{r marginal}
R |>
  filter(target == "none") |>
  group_by(arm, n) |>
  summarise(infit = round(100 * mean(p_infit < ALPHA), 1),
            outfit = round(100 * mean(p_outfit < ALPHA), 1),
            .groups = "drop")
```

## Detection of the misfitting item

```{r det}
R |>
  filter(is_misfit) |>
  group_by(arm, target, loc, n) |>
  summarise(infit = round(100 * mean(f_in), 1),
            outfit = round(100 * mean(f_out), 1),
            union = round(100 * mean(f_any), 1),
            reps = n(), .groups = "drop") |>
  arrange(arm, target, n)
```

```{r fig-det, fig.width = 8, fig.height = 5}
R |>
  filter(is_misfit) |>
  group_by(arm, target, n) |>
  summarise(infit = 100 * mean(f_in), outfit = 100 * mean(f_out),
            .groups = "drop") |>
  pivot_longer(c(infit, outfit), names_to = "statistic", values_to = "det") |>
  ggplot(aes(target, det, colour = statistic, group = statistic)) +
  geom_line() + geom_point(size = 2) +
  facet_grid(arm ~ n, labeller = label_both) +
  scale_x_discrete(limits = c("on", "off"),
                   labels = c("on target", "off target")) +
  ylim(0, 100) +
  labs(x = "Targeting of the misfitting item", y = "Detection (%)",
       colour = "Statistic") +
  theme_minimal(base_size = 10)
```

## False positives on fitting items

Per-item rate among the items that do fit, in the conditions where one item
misfits. This is the cascade check.

```{r fp}
R |>
  filter(target != "none", !is_misfit) |>
  group_by(arm, target, n) |>
  summarise(infit = round(100 * mean(f_in), 2),
            outfit = round(100 * mean(f_out), 2),
            union = round(100 * mean(f_any), 2), .groups = "drop") |>
  arrange(arm, target, n)
```

Direction of the false flags, which in the earlier studies was almost entirely
overfit:

```{r fp-dir}
R |>
  filter(target != "none", !is_misfit, f_in == 1 | f_out == 1) |>
  group_by(arm, target) |>
  summarise(n_flags_infit = sum(f_in),
            overfit_infit = round(100 * mean(infit_msq[f_in == 1] < 1), 1),
            n_flags_outfit = sum(f_out),
            overfit_outfit = round(100 * mean(outfit_msq[f_out == 1] < 1), 1),
            .groups = "drop")
```

## Paired discordance

The decisive table. Both statistics are computed on the same replication from
the same bootstrap, so the marginal detection rates above can be identical while
the two rules disagree on which replications they succeed in. `only_outfit`
counts replications where outfit flags the misfitting item and infit does not,
which is the whole practical case for reporting outfit alongside infit.

`mcnemar_p` is the exact McNemar test on the discordant pairs, which is `NA`
when a cell has none. A cell can have no discordant pairs because both rules
detect everything or because neither detects anything, so read it next to
`both` and `neither` rather than alone.

```{r discord-helper}
# Exact McNemar on the discordant pairs. binom.test() errors on a + b = 0,
# which happens in any cell where the two rules agree in every replication.
exact_mcnemar <- function(a, b) {
  mapply(function(x, y) {
    if (x + y == 0L) NA_real_ else stats::binom.test(x, x + y, 0.5)$p.value
  }, a, b)
}
```

```{r discord}
R |>
  filter(is_misfit) |>
  group_by(arm, target, n) |>
  summarise(both = sum(f_in == 1 & f_out == 1),
            only_infit = sum(f_in == 1 & f_out == 0),
            only_outfit = sum(f_in == 0 & f_out == 1),
            neither = sum(f_in == 0 & f_out == 0),
            reps = n(), .groups = "drop") |>
  mutate(pct_only_outfit = round(100 * only_outfit / reps, 1),
         mcnemar_p = round(exact_mcnemar(only_outfit, only_infit), 4)) |>
  arrange(arm, target, n)
```

Same counts under the null, where every discordant flag is a false positive on
some item:

```{r discord-null}
R |>
  filter(target == "none") |>
  group_by(arm, n, rep) |>
  summarise(i = as.integer(sum(f_in) > 0), o = as.integer(sum(f_out) > 0),
            .groups = "drop") |>
  group_by(arm, n) |>
  summarise(both = sum(i == 1 & o == 1),
            only_infit = sum(i == 1 & o == 0),
            only_outfit = sum(i == 0 & o == 1),
            neither = sum(i == 0 & o == 0), .groups = "drop")
```

## Session

```{r session}
data.frame(
  B = B, alpha = ALPHA, reps = REPS,
  cells = nrow(s$grid), elapsed_min = round(s$elapsed_min, 1),
  easyRasch2 = as.character(utils::packageVersion("easyRasch2")),
  R = paste(R.version$major, R.version$minor, sep = ".")
)
```
